New lower bounds for Schur and weak Schur numbers

### 2021-10-28 article/weakschur.tex

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\) and
a sum-frree \(k\)-partition of \([\![1,p]\!]\) then there is a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right 
\rceil + 1)+q]\!]}\) into \(n+k\) weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

@@

Let \((p, k), (q, n) \in (\mathbb{N}^*)^2\) such that there are a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets. Let \(a \in \mathbb{N}\) with \(a > q\)
and let us try to build a coloring of \([\![1, ap + q]\!]\) into \(n + k\) weakly sum-free subsets. Let
\(l = a - q - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = q - r\).

First, we put the integers of \([\![1, ap + q]\!]\) in the following table (with \(a\) columns and \(p + 1\) lines)
and divide it into 3 blocks (the columns are numbered from \(-l\) to \(+q\)).

@@

\begin{definition}
Let \( n \in \mathbb{N}^*\). We define:
\[
\WS^+(n)=\max_{b\in \mathbb{N}^*} \WS_b^+(n).
\]
\end{definition}
The following proposition briefly shows how \(\WS^+(n)\) compares to weak Schur numbers.
\begin{proposition}
Let \(n \in [\![2, +\infty]\!]\). Then:
\[
\frac{3}{2} \WS (n-1)+1 \leqslant \WS^+(n) \leqslant \WS (n).
\]
\end{proposition}

@@

Theorem \ref{thm:Stemp} describes a manner to extend a S-template into a larger Schur partitions using other Schur 
partitions. Likewise, a WS-template can be extended into larger weak Schur partitions by using Schur partitions. This 
is the object of the following theorem.

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there are a partition of \([\![1,p]\!]\)
into \(k\) sum-free subsets and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there is a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

@@

Inequality (\ref{WS(n+3)}), found with a SAT solver \cite{Lingeling2017}, cannot be improved (with this 
definition of WS-template). It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We expect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions.

>>>>

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\)
and a sum-free \(k\)-partition of \([\![1,p]\!]\) then there is a partition of
\({[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]}\) into \(n+k\) weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

@@

Let \((p, k), (q, n) \in (\mathbb{N}^*)^2\) such that there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\) 
and a sum-free \(k\)-partition of \([\![1,p]\!]\). Let \(a \in \mathbb{N}\) with \(a > q\)
and let us try to build a coloring of \([\![1, ap + q]\!]\) into \(n + k\) weakly sum-free subsets. Let
\(l = a - q - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = q - r\).

First, we construct the following table with \(a\) columns and \(p+1\) rows containing the integers from \([\![1, ap + q]\!]\). 
The columns are numbered from \(-l\) to \(+q\). This table is made of three different blocks.

@@

\begin{definition}
Let \( n \in \mathbb{N}^*\). We define:
\[
\WS^+(n)=\max_{b\in \mathbb{N}^*} \WS_b^+(n).
\]
\end{definition}
The following proposition shows how \(\WS^+(n)\) compares to weak Schur numbers.
\begin{proposition}
Let \(n \in [\![2, +\infty]\!]\). Then:
\[
\frac{3}{2} \WS (n-1)+1 \leqslant \WS^+(n) \leqslant \WS (n).
\]
\end{proposition}

@@

Theorem \ref{thm:Stemp} describes a manner to extend an S-template into a larger Schur partitions using other Schur 
partitions. Likewise, a WS-template can be extended into larger weak Schur partitions by using Schur partitions. This 
is the object of the following theorem.

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there are a sum-free \(k\)-partition of
\([\![1,p]\!]\) and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there is a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

@@

Inequality (\ref{WS(n+3)}), found with a SAT solver \cite{Lingeling2017}, cannot be improved (with this 
definition of WS-template). It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining an S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We expect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions.
